Refinement for open automata

### 2021-05-03 2021-Refinement/report.tex
Parellel non-locking composition proof begun

% possible examples for refinement: + <= |, ; <= |
% transitions between sections
% Re-read what refinement are used for, LOTOS
% read CCSL and BIP

@@

\section{Introduction}
Open automata are used to give an interpretation for open pNets.
% TODO: expand

\section{Notations}
Notations will be defined with the operator \(\defnotation\) and names are given with the operator \(\defobject\) as follows:
\begin{align*}
	notation\_with\_variables & \defnotation notated\_object\_using\_the\_variables \\
	name & \defobject fully\_defined\_mathematical\_object
\end{align*}

@@

\section{Open Automaton}
To define an open automaton we need some preliminary definitions.
\begin{defi}[Expression algebra, Action algebra, Formulae, Terms]
An expression algebra \(E\) is a disjoint union \(E \defobject \terms \uplus \actions \uplus \formulae\) of the terms, the actions and the formulae.

@@

\begin{figure}
\centering
\input{enable_state.tex}
\vrule
\input{enable_var.tex}
\caption{Enable operator implementation with open automata, on the left state oriented, on the right variable oriented}
\label{fig:enable}
\end{figure}
To illustrate these definitions let's look at two implementations (figure \ref{fig:enable}) of the LOTOS\marginpar{cite something here} operator enable in the open automata model.
The enable operator runs its the left hand side agent until it chooses to finish, at which point it produces an action \(\delta\) that is synchronised with the first action of the right hand side agent which must be \(\act{acc}\), then only the latter agent runs.
During the synchronised action \Quentin{data can be exchanged}{I need to read a little bit more on that; is that a return code?}.
\begin{exi}[Enable, state-oriented]
To draw automata, the usual circle for states and simple arrows for transitions convention is used.
The initial state is indicated by a double circle.
States names are indicated inside the circles and transitions labels are drawn near their corresponding arrow.

@@

This automaton is an implementation of the enable operator because it begins in the state \(L\) where it allows any transition from its hole \(l\) which is not \(\delta\) then synchronises its holes on the same data, effectively allowing a data exchange when it goes into state \(R\), and finally allows any transition from its hole \(r\).
Only the synchronisation is hidden to the exterior by sending a \(\tau\) which is usually a silent action always allowed, unsynchronised with other holes and passed unmodified.
Well here \(\tau\) is not always allowed from the holes in the state \(L\) the hole \(r\) cannot emmit it for the sake of simplifying the first example of open automaton.
\end{exi}
\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which has already been proven equivalent in previous work by ...\marginpar{cite sth here}.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are like previously all the transitions generated by
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) is a hole name and also a closed term, but not a variable.
The same applies for \(l\).
\(v\) is a variable name for the automaton and a variable in the guard, so its value is not substituted.
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the substitution \(v \mapsto l\)).

@@

Finally we can define some utilitary functions:
\begin{defi}[Guard, Out-transition]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(s\) be one of its states.
\(\fOT{s}\) are called the out-transitions of \(s\).
\begin{align*}
	\fguard{\OTg} & \defnotation g &
	\fOT{s} & \defnotation \mbrc{\OTg \in T}
\end{align*}
\[ \fvars{\OTg} \defnotation \mpar{\fvars{\alpha} \cup \fvars{g} \cup \bigcup_{j \in J'} \fvars{\beta_j} \cup \bigcup_{v \in V} \fvars{\psi\mpar{v}}} \setminus V \]
\end{defi}
From this point open automata and open transitions will be called automata and transitions for simplicity.

\section{Semantic and composition of Open Automata}
The interpretation of open automata is a partially specified automaton, where the partially specification comming from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work\marginpar{cite 2007.10770} but never completely formalised on open automata.
The below definition of composition is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
The component we choose to compose in the holes are in figure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(C\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Full.tex}
\caption{The full traffic lights system}
\label{fig:tlf}
\end{figure}
The composition of the specification on figure \ref{fig:tls} and the agents on figure \ref{fig:tlh} is the automaton on figure \ref{fig:tlf}.
Unreachable states and transition with unsatisfiable guards are not drawn.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
Also satisfiable guards were replaced with simplified equivalent.

@@

Let's look at the \(\tau\) transition from the state \(R\) for a more in depth example on how composition and value passing work.
The transition in the specification is \nmm{\OT{R}{R}{\tau}{\mbrc{R \mapsto \act{set}\mpar{x}, C \mapsto \theta\mpar{x}}}{\top}{\mbrc{}}}.
The composition makes 12 states from the \(R\) state, these states are in \(\mbrc{R} \times \mdbrk{1; 6} \times \mbrc{S, C}\), however only \(\mpar{R, 1, S}\) and \(\mpar{R, 2, C}\) are reachable so we will only consider these two.
In the figure they are noted \(RS\) and \(RC\).
The holes \(R\) and \(C\) are filles at the same time so the \(\tau\) transition must synchronise with a transition from both the register and the controller.

In the state \(RC\) the composition makes 2 transition from the \(\tau\) transition with a false guard, let's look at one of them.
\nmm{\OT{RC}{\mpar{R, 3, C}}{\tau}{\mbrc{}}{\top \wedge \top \wedge c < t \wedge \act{set}\mpar{x} = \act{tick} \wedge \theta\mpar{x} = \delta\mpar{x_C}}{\mbrc{c \gets c + 1}}} is made with \nmm{\OT{2}{3}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller and \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
Obviously the guard is false so the it is equivalent to have no transition.
Also the guard is long and can be simplified.
In this article the guard and transition are always simplified except for the sake of illustrating a concept that is lost in the simplification as it is the case here.
We can see that the \(x\) variables from the specification, the controller and the register are in conlict.
To resolve the conflict we renamed the variables.

In the state \(RS\) the composition makes 1 transition from the \(\tau\) transition; this transition is described before but let's look at the details of the composition.
% TODO: explain value passing then say that in the register \set{t} wouldn't have worked
\end{exi}

\section{What is a refinement for Open Automata}
There are several properties that we may want from a refinement relation.
Depending on these properties, several kind of refinement may be used.
For example if we are interested in being able to produce the same sequence of action we may want to use trace set inclusion as a refinement.
Here the main properties that we want are related to composition and action refinement.
A suitable kind of refinement relation for this kind of properties is simulation refinement:
A bisimulation has already been proposed for open automata \marginpar{cite an article here} that is compatible with composition.

% THIS WILL BE RE-WRITTEN LATER
Other kind of refinement relation that \Quentin{I may explore}{This is a note for possible path, some may be explored, some may not.} are control refinement/hole refinement\footnote{No good formulation atm, the idea is that \(a \leq b \defnotation sth \wedge J_a \setminus J_b \leq_{ctrl} J_b \setminus J_a\), sth is probably a clause to ensure that they behave the same without holes involved.}, (meet semi)lattice refinement\footnote{meet = handshake, \(a \leq b \defnotation a = a || b\) with sync holes, I think this will be equivalent to hole-identical sim, (join=non-det choice?)}, weak-simulation refinement\footnote{Weak variation for each interesting simulation refinement}, composition-correct refinement\footnote{Basically a stricter variant of simulation-refinement where simulation has to take place the other way for some transitions (no different holes?) in order to be correct wrt composition}.

@@

% LATER: livelock and tau-stuttering
% \item[Livelock reduction:] If \(a \leq b\) then \(a\) does not introduces new livelocks, noted \(a \preceq b\) in the following; % TODO: Find a better place to introduce a better definition of livelock reduction.
% \defitem \textbf{\(\tau\)-stuttering} iff . % TODO

@@

To do that we impose that the initial state is reachable in the initial valuation and that reachability is preserved by taking valid transitions.
This effectively makes reachability account for all paths, and to not account for impossible path the strongest predicate is used.
Another way to understand this predicate is that it is the least fixpoint of the union of valuations that contains the initial valuation.
\begin{defi}[Non-locking composition, intuitive definition]
The composition resulting in \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if:
\defitem From reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in T, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}

\begin{exi}
\begin{figure}
\input{randomtick_clock.tex}
\vrule
\input{register_anytick.tex}
\caption{On the left: A clock which imposes a tick; On the right: A modified version of the register at figure \ref{fig:tlh}}
\label{fig:anytick}
\end{figure}
% TODO HERE: explain
\begin{figure}
\input{Composition_deadlock.tex}
\vrule
\input{Composition_nolock.tex}
\caption{On the left: Deadlock introduced by composition; On the right: No deadlock introduced by composition}
\label{fig:deadlock}
\end{figure}
% TODO: explain
\end{exi}

@@

\begin{defi}[Deadlock reduction, working definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}
Proof is given in the appendix. % TODO: ref

\begin{defi}[Non-locking composition, working definition]
The composition \(A \defobject A_p\subst{A_c}{k}\) is a non-locking composition if:
\[ \forall s \in S, \forall \sigma: V \to \values, \sigma \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_p \in \fOT{\pi_{S_p}\mpar{s}}} \fguard{t_p} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \]
\end{defi}
\begin{lem}
The intuitive and working definition of non-locking composition are equivalent.
\end{lem}
Proof is given in the appendix. % TODO: ref and proof in apendix

% TODO: sufficient condition

From this point and in the previous definitions, composition will only refer to correct composition.

>>>>

% possible examples for refinement: + <= |, ; <= |
% transitions between sections

@@

\section{Introduction}
The model of open automata is used to give the semantic of open pNets \cite{2007.10770}.
Open pNets are not petri nets but an intermediate representation used in the VerCors project to perform verification and model checking \cite{henrio:01252323}.
Open pNets with the open automaton semantic and equivalence relation introduced in previous work \cite{2007.10770} were sucessfully used to verify BIP and GCM \cite{qin:01823507, ameurboulifa:01526055}.
However equivalences relations are not sufficient to perform model checking, refinement relations are also used.
For instance an example of open pNet where equivalence is not sufficient was encountered in ...\marginpar{Paper where the motivating example came up}.

In this article we introduce a refinement relation for open automata. % TODO: expand

We begin in section \ref{sec:notations} by giving some notations used throughout this paper.
Then there are several sections dedicated to define all the interesting objects and properties:
In section \ref{sec:def} we give a clean and up to date definition of open automata.
In section \ref{sec:semcompo} we recall the semantic for open automata and their composition without requiring the use of open pNets.
In section \ref{sec:proofelts} we adapt the standard properties related to refinement relation from LTS to open automata.
After that the section \ref{sec:prelref} is dedicated to progressively build a refinement relation.
Finally the section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.

\section{Notations}\label{sec:notations}
Notations will be defined with the operator \(\defnotation\) and names are given with the operator \(\defobject\) as follows:
\begin{align*}
	notation\_with\_variables & \defnotation notated\_object\_using\_the\_variables \\
	name & \defobject fully\_defined\_mathematical\_object
\end{align*}

@@

\section{Open Automaton}\label{sec:def}
To define an open automaton we need some preliminary definitions.
\begin{defi}[Expression algebra, Action algebra, Formulae, Terms]
An expression algebra \(E\) is a disjoint union \(E \defobject \terms \uplus \actions \uplus \formulae\) of the terms, the actions and the formulae.

@@

\begin{noti}[Bisimilarity, FH-bisimulation]
The bisimilarity relation on open automata is noted \(\cong\) and the FH-bisimulation \cite{henrio:01055091} is noted \(\overset{FH}=\).
\end{noti}
Note that FH-bisimilarity implies bisimilarity.

To illustrate these definitions let's look at two implementations (figure \ref{fig:enable}) of the LOTOS\marginpar{cite something here} operator enable in the open automata model.
The enable operator runs its the left hand side agent until it chooses to finish, at which point it produces an action \(\delta\) that is synchronised with the first action of the right hand side agent which must be \(\act{acc}\), then only the latter agent runs.
During the synchronised action \Quentin{data can be exchanged}{I need to read a little bit more on that; is that a return code?}.
\begin{exi}[Enable, state-oriented]
\begin{figure}
\centering
\input{enable_state.tex}
\vrule
\input{enable_var.tex}
\caption{Enable operator implementation with open automata, on the left state oriented, on the right variable oriented}
\label{fig:enable}
\end{figure}
To draw automata, the usual circle for states and simple arrows for transitions convention is used.
The initial state is indicated by a double circle.
States names are indicated inside the circles and transitions labels are drawn near their corresponding arrow.

@@

This automaton is an implementation of the enable operator because it begins in the state \(L\) where it allows any transition from its hole \(l\) which is not \(\delta\) then synchronises its holes on the same data, effectively allowing a data exchange when it goes into state \(R\), and finally allows any transition from its hole \(r\).
Only the synchronisation is hidden to the exterior by sending a \(\tau\) which is usually a silent action always allowed, unsynchronised with other holes and passed unmodified.
Well here \(\tau\) is not always allowed from the holes in the state \(L\) the hole \(r\) cannot emmit it for the sake of simplifying the first example of open automaton.
\end{exi}
\begin{exi}[Enable, variable-oriented]
The automata drawn on the right side of the figure is an alternative implementation of the enable operator which is FH-bisimilar.
The drawing introduces the double arrows which is used to denote the initial valuation.
The automata drawn is \(\OA{\mbrc{\_}}{\_}{\mbrc{v}}{l^\mbrc{v}}{\mbrc{l, r}}{T}\) where transitions in \(T\) are like previously all the transitions generated by
\begin{align*}
	\OT{\_}{\_}{x}{\mbrc{l \mapsto x}}{v = l \wedge x \neq \delta}{\mbrc{}} &&
	\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \delta\mpar{x}, r \mapsto \act{acc}\mpar{y}}}{v = l \wedge x = y}{\mbrc{v \gets r}} &&
	\OT{\_}{\_}{x}{\mbrc{r \mapsto x}}{v = r}{\mbrc{}}
\end{align*}
Note that in this example \(r\) is a hole name and also a closed term, but not a variable.
The same applies for \(l\).
\(v\) is a variable name for the automaton and a variable in the guard, so its value is not substituted.
For instance \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{v = l \wedge \tau \neq \delta}{\mbrc{}}} is in \(T\) but not \nmm{\OT{\_}{\_}{\tau}{\mbrc{l \mapsto \tau}}{l = l \wedge \tau \neq \delta}{\mbrc{}}} (result of the substitution \(v \mapsto l\)).

@@

Finally we can define some utilitary functions:
\begin{defi}[Guard, Out-transition, Transition variables]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(s\) be one of its states.
\(\fOT{s}\) are called the out-transitions of \(s\).
\begin{align*}
	\fguard{\OTg} & \defnotation g &
	\fOT{s} & \defnotation \mbrc{\OTg \in T}
\end{align*}
\[ \fvars{\OTg} \defnotation \mpar{\fvars{\alpha} \cup \fvars{g} \cup \bigcup_{j \in J'} \fvars{\beta_j} \cup \bigcup_{v \in V} \fvars{\psi\mpar{v}}} \setminus V \]
\end{defi}
From this point open automata and open transitions will be called automata and transitions for simplicity.

\section{Semantic and composition of Open Automata}\label{sec:semcompo}
The interpretation of open automata is a partially specified automaton, where the partially specification comming from the holes.
This specification can be completed by composing automata into holes so it is natural to take a look at the composition process.
This was already defined on pNets in previous work \cite{henrio:01299562} but never completely formalised on open automata.
The below definition of composition is a direct translation of composition on pNets for open automata.
\begin{defi}[Composition of open automata]
The composition of \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\) in the hole \(k \in J_p\) of \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) is
\begin{align*}
	A_p\subst{A_c}{k} \defnotation & \OA{S_p \times S_c}{\mpar{s_{0p}, s_{0c}}}{V_p \uplus V_c}{\sigma_{0p} \uplus \sigma_{0c}}{J_c \uplus J_p \setminus \mbrc{k}}{T} \\
	\text{With } T \defobject & \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s'_c}}{\alpha_p}{\beta_j^{j \in J'_c \uplus J'_p \setminus \mbrc{k}}}{g_p \wedge g_c \wedge \alpha_c = \beta_k}{\psi_p \uplus \psi_c}}{\OTx{p}{}{}{p} \in T_p, \OTx{c}{}{}{c} \in T_c} \\
	& \cup \mset{\OT{\mpar{s_p, s_c}}{\mpar{s'_p, s_c}}{\alpha_p}{\beta_j^{j \in J'_p}}{g_p}{\psi_p}}{\OTx{p}{}{}{p}, k \notin J', s_c \in S_c}
\end{align*}
Similarly to expression substitution, the parallel composition is noted \(A\psubst{A_j^{j \in J}}\).
\end{defi}
The actions emitted when \(A_c\) makes a transition is sychronised with the action of the hole \(k\) in transitions of \(A_p\) which have it as a hole actions (first transition set, \(\alpha_c = \beta_k\)).
No transition of \(A_c\) is performed when a transition that do not refer to the hole \(k\) is performed in \(A_p\) (second transition set, \(k \notin J'\)).
The composition may look like a handshake, with both automata running in parallel (product of states and joint variables) but it is actually not symmetric.
\begin{exi}
This example is derived and adapted from a traffic light controller in a collection of examples for pNets.
It is supposed to be a single traffic light.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Spec.tex}
\caption{The specification of a traffic light system}
\label{fig:tls}
\end{figure}
\begin{figure}
\centering
\input{Traffic_Lights_Controller.tex}
\vrule
\input{Traffic_Lights_Register.tex}
\caption{On the left: The controller agent; On the right: The counter agent}
\label{fig:tlh}
\end{figure}
The component we choose to compose in the holes are in figure \ref{fig:tlh}.
On the left is the controller agent which will be composed in the hole \(C\).
Its role is to decide the duration before switching the lights.
This implementation is simply a constant time for each color.

@@

The automaton on figure \ref{fig:tlf} is bisimilar to the composition of the specification on figure \ref{fig:tls} and the agents on figure \ref{fig:tlh}.
The result of the composition has 36 states (\(3 \times 2 \times 6\)) but only 6 are reachable from the initial configuration.
Also satisfiable guards were replaced with simplified equivalent.

@@

\begin{figure}
\centering
\input{Traffic_Lights_Full.tex}
\caption{The full traffic lights system}
\label{fig:tlf}
\end{figure}
Let's look at the \(\tau\) transition from the state \(R\) for a more in depth example on how composition and value passing work.
The transition in the specification is \nmm{\OT{R}{R}{\tau}{\mbrc{R \mapsto \act{set}\mpar{x}, C \mapsto \theta\mpar{x}}}{\top}{\mbrc{}}}.
The composition makes 12 states from the \(R\) state, these states are in \(\mbrc{R} \times \mdbrk{1; 6} \times \mbrc{S, C}\), however only \(\mpar{R, 1, S}\) and \(\mpar{R, 2, C}\) are reachable so we will only consider these two.
In the figure they are noted \(RS\) and \(RC\).
The holes \(R\) and \(C\) are filles at the same time so the \(\tau\) transition must synchronise with a transition from both the register and the controller.

In the state \(RC\) the composition makes 2 transition from the \(\tau\) transition but both have a false guard; let's look at one of them.
\nmm{\OT{RC}{\mpar{R, 3, C}}{\tau}{\mbrc{}}{\top \wedge \top \wedge c < t \wedge \act{set}\mpar{x} = \act{tick} \wedge \theta\mpar{x} = \delta\mpar{x_C}}{\mbrc{c \gets c + 1}}} is made with \nmm{\OT{2}{3}{\delta\mpar{x}}{\mbrc{}}{\top}{\mbrc{}}} from the controller and \nmm{\OT{C}{C}{\act{tick}}{\mbrc{}}{c < t}{\mbrc{c \gets c + 1}}} from the register.
Obviously the guard is false so the it is equivalent to have no transition.
Also the guard is long and can be simplified.
In this article the guard and transition are always simplified except for the sake of illustrating a concept that is lost in the simplification as it is the case here.
We can see that the \(x\) variables from the specification, the controller and the register are in conlict.
To resolve the conflict we renamed the variables.

In the state \(RS\) the composition makes 1 transition from the \(\tau\) transition.
This transition is described before but let's look at the details of the composition.
The obtained guard is \(\top \wedge \top \wedge \top \wedge \theta\mpar{x} = \theta\mpar{17} \wedge \act{set}\mpar{x} = \act{set}\mpar{x_R}\) and the obtained new values of the variables is \(\mbrc{t \gets x_R, c \gets 0}\).
We can see here how value passing work: the only possible value of \(x\) and \(x_R\) making the guard hold is \(17\), this value has been synchronised from the controller to \(x\) then to the second occurence of \(x\) to the distinct \(x\) of the register.
This guard and value update can be simplified to the one shown on the drawing.
Note that \nmm{\OT{S}{C}{\act{set}\mpar{t}}{\mbrc{}}{\top}{\mbrc{c \gets 0}}} wouldn't have worked as a transition for the controller because \(t\) is fixed before the transition and its value cannot be set outside of the variable update, hence the use of a local variable \(x\).
\end{exi}

\section{What is a refinement for Open Automata}\label{sec:proofelts}
There are several properties that we may want from a refinement relation.
Depending on these properties, several kind of refinement may be used.
For example if we are interested in being able to produce the same sequence of action we may want to use trace set inclusion as a refinement.
Here the main properties that we want are related to composition and action refinement.
Because of composition, a relation as strong as simulation refinement must be used as observed in the articles about FH-bisimulation \cite{henrio:01055091}.

% LATER
% Other kind of refinement relation that \Quentin{I may explore}{This is a note for possible path, some may be explored, some may not.} are control refinement/hole refinement\footnote{No good formulation atm, the idea is that \(a \leq b \defnotation sth \wedge J_a \setminus J_b \leq_{ctrl} J_b \setminus J_a\), sth is probably a clause to ensure that they behave the same without holes involved.}, (meet semi)lattice refinement\footnote{meet = handshake, \(a \leq b \defnotation a = a || b\) with sync holes, I think this will be equivalent to hole-identical sim, (join=non-det choice?)}, weak-simulation refinement\footnote{Weak variation for each interesting simulation refinement}, composition-correct refinement\footnote{Basically a stricter variant of simulation-refinement where simulation has to take place the other way for some transitions (no different holes?) in order to be correct wrt composition}.

@@

% LATER: livelock and tau-stuttering
% \item[Livelock reduction:] If \(a \leq b\) then \(a\) does not introduces new livelocks, noted \(a \preceq b\) in the following; % Find a better place to introduce a better definition of livelock reduction.
% \defitem \textbf{\(\tau\)-stuttering} iff .

@@

To do that we impose that the initial state is reachable in the initial valuation and that reachability is preserved by taking valid transitions.
This effectively makes reachability account for all paths, and to not account for impossible path the strongest predicate is used.
Another way of understanding this predicate is that it is the least fixpoint of the union of valuations that contains the initial valuation:
\begin{align*}
	f\mpar{p} = s' \mapsto p\mpar{s'} \vee \bigsymb{\bigvee_\subbox{\OTg \,\in\, T}} p\mpar{s}\psubst{\psi} \wedge g &&
	\reach{A} = \lim_n f^n\mpar{s \mapsto \choice{\sigma_0 & s = s_0 \\ \bot & \text{otherwise}}}
\end{align*}

\begin{defi}[Non-locking composition, intuitive definition]
The composition resulting in \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if:
\defitem From reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in T, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}

To illustrate the introduction of a deadlock in a composition and the definition of non-locking composition, let's look at an example.
\begin{exi}[Deadlocks introduced by composition]
% TODO HERE: read CCSL & BIP
\begin{figure}
\input{randomtick_clock.tex}
\vrule
\input{register_anytick.tex}
\caption{On the left: A clock which imposes a tick; On the right: A modified version of the register at figure \ref{fig:tlh}}
\label{fig:anytick}
\end{figure}
% TODO: explain
\begin{figure}
\input{Composition_deadlock.tex}
\vrule
\input{Composition_nolock.tex}
\caption{On the left: Deadlocks introduced by composition; On the right: No deadlock introduced by composition}
\label{fig:deadlock}
\end{figure}
% TODO: explain
\end{exi}

@@

\begin{defi}[Deadlock reduction, working definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \to \values, \sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}

\begin{defi}[Non-locking composition, working definition]
The composition \(A \defobject A_p\subst{A_c}{k}\) is a non-locking composition if:
\[ \forall s \in S, \forall \sigma: V \to \values, \sigma \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_p \in \fOT{\pi_{S_p}\mpar{s}}} \fguard{t_p} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \]
\end{defi}
\begin{lem}
The intuitive and working definition of non-locking composition are equivalent.
\end{lem}
Proof of both lemma are given in appendix \ref{apx:lemeqd}.

% TODO?: sufficient condition
